---
title: "Project_Version_1.1"
author: "Hanwen Zhang"
date: "12/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import libraries which will be used in this project
Import the libraries which will be used in this project
```{r}
library(tidyverse)
library(ggthemes)
library(corrplot)
library(GGally)
library(caret)
library(devtools)
library(data.world)
library(DT)
library(lubridate)
library(factoextra)
library(ggplot2)
library(gapminder)
options(knitr.table.format = "html")
```

# Import datasets
```{r}
death_data <- read.csv("https://query.data.world/s/hj7q35azsakdpkngaqx6npp7z7eobo", fileEncoding="latin1", header = TRUE)
incd_data <- read.csv("https://query.data.world/s/fxovg6so5pfdndl2sz7xjso3l7y3u6", fileEncoding="latin1", header = TRUE)
```

# Data preprocessing
In both of the data sets, replace the value * to null value NA
```{r}
death_data[death_data == "*"] <- NA
death_data[death_data == "**"] <- NA
incd_data[incd_data == "*" ] <- NA
incd_data[incd_data == " *"] <- NA
incd_data[incd_data == "* "] <- NA

```

# creat functions
```{r}
# Bar plot for Missing Values of a data set
column_nas = function(dataframe) {
        # Extract the proportion of null values in the data frame
        na_vals = sapply(dataframe , function(x) mean(is.na(x)))
        # Generate a new data frame named nas 
        nas = data.frame(column_names = names(na_vals) , na_percentage = as.vector(na_vals))
        # Plot the proportion of the null values of each data set's variable
        ggplot(nas , aes(x = column_names , y = na_percentage , label = paste(as.character(round(na_percentage * 100 , 1)) , "%"))) +
                geom_col(fill = "lightblue") + xlab('Names of columns') + ylab("NA's Percentage") +
                labs(title = "Column NA's Percentage") + geom_text(color = 'darkblue') + theme_igray() + coord_flip()
}

# Plot for every variable. Bar plot for categorical and density histogram for numerical
plot.var = function(var) {
        
        if(class(var) == 'numeric' | class(var) == 'integer') {
                
                ggplot(data.frame(var) , aes(var)) + geom_density(color = 'blue' , size = 1 , na.rm = T) + 
                        theme_bw() + theme(axis.text.x = element_text(size = 15)) + 
                        geom_vline(xintercept = mean(var))
                
        } else if(class(var) == 'factor') {
                
                ggplot(data.frame(var) , aes(x = var , y = ..count..)) + 
                        geom_bar(fill = 'blue' , width = 0.5) + coord_flip()
                
        } else {
                print('Variable NOT numeric or integer or factor')
        }
        
}
```

# Plot the proportion of the null values of each data set's variable
```{r}
column_nas(death_data)
column_nas(incd_data)
```
Remove "Met.Objective.of.45.5...1.", "Average.Deaths.per.Year", 

# Column names of death_data
```{r}
colnames(death_data)
```

# Column names of incd_data
```{r}
colnames(incd_data)
```

# Extract the data which will be used
```{r}
death_data = death_data[ , c(1 , 2 , 4 , 7)]
incd_data = incd_data[ , c(1, 2, 3, 6)]
```

# Changing variable name
```{r}
colnames(death_data)[3:4] = c("death_rate" , "ann_avg_deaths")
colnames(incd_data)[3:4] = c("incidence_rate" , "ann_avg_incidence")
```

# Merge two data sets by FIPS
```{r}
total_set = merge(x = death_data , y = incd_data , by = 'FIPS' , all.x = TRUE)
```

# Remove 'county.y' variable
```{r}
total_set = total_set[ , c(1, 2, 3, 4, 6, 7)]
```

# Rename county.x to county
```{r}
colnames(total_set)[2] = "county"
```

# Change incidence_rate and ann_avg_incidence to numeric
```{r}
total_set$incidence_rate = as.numeric(total_set$incidence_rate)
total_set$ann_avg_incidence = as.numeric(total_set$ann_avg_incidence)
total_set$death_rate = as.numeric(total_set$death_rate)
#total_set$ann_avg_deaths = as.numeric(total_set$ann_avg_deaths)

datatable(total_set , options = list(pageLength = 5))
```

# Summary Statistics for death
```{r}
summary(total_set$death_rate)
```

# Summary Statistics for death
```{r}
summary(total_set$incidence_rate)
```

# Gather data to a new dataset
```{r}
gathered_set = gather(total_set , 'Death.or.Incidence' , 'Rate' , c(3 , 5))
```

# Fix levels on Death.or.Incidence column
```{r}
gathered_set$Death.or.Incidence = ifelse(gathered_set$Death.or.Incidence == 'death_rate' , 'Death Rate' , 'Incidence Rate')
```

# Density Plot for the two rates
```{r}
ggplot(gathered_set[complete.cases(gathered_set) , ] , aes(Rate , fill = Death.or.Incidence)) + geom_density(alpha = 0.5) +
        xlim(c(0 , 150)) + theme_economist()
```

```{r}
#rm(gathered_set)
```

```{r}
#devtools::install_github("deleetdk/USA.county.data")
```

```{r}
usa_data <- USA.county.data::USA_county_data
```

```{r}
# Cut the 161 variables to 4 pieces of 40 , 40 , 40 and 41. 
a = list(1:40 , 41:80 , 81:120 , 121:161)
```

```{r}
str(usa_data[ , a[[1]]])
```

```{r}
column_nas(usa_data[ , a[[1]]])
```

```{r}
# New data set with these variables
new = usa_data[ , c(1, 72:75, 77, 119:120, 86, 98:102, 84:85, 88, 90, 115:117)]
datatable(new)
```

```{r}
column_nas(new)
```

Since there are too many missing values in Sexually.transmitted.infections
Remove the column
```{r}
new = new[ , -(21)]
```

Remove the rows which contained missing values
And make it a new data set
```{r}
missing_values_rows = complete.cases(new)
new = new[missing_values_rows , ]
rm(missing_values_rows)
cat("New dimensions after removing NA's" , "\nRows:" , dim(new)[1] , "\nColumns:" , dim(new)[2])
```

Observe the predictors and transforming

Observe the education predictors
```{r}
ggpairs(new[ , c(2 , 3 , 4 , 5)] , aes(alpha = 0.2 , color = 'darkblue')) + theme_bw()
```

According to the plots above, we can find that 
Less.Than.High.School with At.Least.High.School.Diploma 
and At.Least.Bachelor.s.Degree with Graduate.Degree variables 
are highly correlated as expected so there is no need to keep all of them. 
For that reason we'll remove Less.Than.High.School and Graduate.Degree columns.

```{r}
# Remove variables
new = select(new , -Less.Than.High.School , -Graduate.Degree)
```

Population Predictors
```{r}
ggpairs(new[ , 8:12] , aes(alpha = 0.2 , color = 'darkblue'))
```

Poverty predictors
```{r}
ggpairs(new[ , 13:16] , aes(alpha = 0.2 , color = 'darkblue'))
```
Incident and death rates that we use are age adjusted and given a high correlation between Poverty.Rate.below.federal.poverty.threshold and Child.Poverty.living.in.families.below.the.poverty.line 
we'll only keep the Poverty.Rate.below.federal.poverty.threshold column which seems to be representative for poverty rates.

```{r}
# Remove variables
new = new[ , -c(13:14, 16)]
```

Health predictors
```{r}
ggpairs(new[ , 14:15] , aes(alpha = 0.2 , color = 'darkblue'))
```
Given a high correlation between obesity and diabetes (cor = 0.718) 
we will remove Diabetes variable because it is more specific and obesity 
represents a bigger sample of the population.

```{r}
new = new[ , -15]
```

```{r}
colnames(new)[1] = c("FIPS")
```

Merge total_set and new
```{r}
# Merge datasets
final_set = merge(x = total_set , y = new , by = 'FIPS' , all.x = TRUE)

# Remove first row which is for United States as a whole data and Incidence Rate columns
final_set = final_set[-1 , -c(4:6)]

# Remove rows with missing values
final_set = final_set[complete.cases(final_set) , ]

datatable(final_set , options = list(pageLength = 5))
```

Full Analysis

1) Correlation of the variables with Death Rate
```{r}
cors = cor(final_set[ , -c(1,2)])[-1 , 'death_rate']

cors = data.frame(names = names(cors) , cors = cors)
cors$names = as.character(cors$names)
cors$col = ifelse(cors$cors > 0 , 'Pos cor' , 'Neg cor')

ggplot(cors , aes(x = reorder(names , cors) , y = cors , fill = col)) + geom_col() +
        coord_flip() + lims(y = c(-1 , 1)) +
        labs(x = 'Variables Names' , y = 'Correlations' , title = 'Correlations with Death Rate Variable') + 
        theme(legend.position = 'none')
```

```{r}
# rm(cors)
```

2) Correlation plot of the final set
```{r}
corrplot(cor(final_set[ , -c(1:2)]) , type = 'lower' , method = 'number' , cl.srt = 90)
```

3) PCA - Biplot
```{r}
res.pca <- prcomp(final_set[ , -c(1:2)], scale = TRUE)

fviz_pca_biplot(res.pca, col="darkred", col.ind = final_set$death_rate ,
                geom = "point", repel=TRUE , legent.title = 'death_rate') +
        ylim(c(-7 , 2.5)) + scale_color_gradient2(low = 'white' , high = 'darkgreen')
```

4) Regression Model
We ’ll create a multiple linear regression with our dependent variable be death_rate
```{r}
# Regression model
model = lm(death_rate ~ . -FIPS -county , data = final_set)

# Summary of the model
summary(model)
```
Our model expains 50.65 % of the total variance with a standard error of 9.863.

5) Variables Importance
Let’s see the importance of our independent variables.
```{r}
imp = varImp(model)
imp = data.frame(names = row.names.data.frame(imp) , overall = imp$Overall)

ggplot(imp , aes(x = reorder(names , overall) , y = overall , label = round(overall , 1))) + 
        geom_point(fill = 'black' , size = 12) + coord_flip() +
        geom_text(color = 'white' , size = 5) + 
        labs(x = 'Variables' , y = 'Importance')
```

```{r}
# rm(imp)
```

6) Bias of the estimator
```{r}
norm = data.frame(value = rnorm(100000 , mean = 0 , sd = sd(model$residuals)) , norm.res = 'Normal dist.')
resids = data.frame(value = model$residuals , norm.res = 'Residuals')

a = bind_rows(norm , resids)

ggplot(a , aes(x = value , fill = norm.res)) + geom_density(alpha = 0.6) + 
        labs(title = 'Residuals density plot vs Normal distribution' , subtitle = 'Residual standard error: 9.863' )
```

```{r}
# rm(a , norm , resids , resids)
```

7) Actual and Fitted values
```{r}
ggplot(data.frame(actual = final_set$death_rate , fitted = model$fitted.values) , aes(x=actual , y=fitted)) + 
        geom_jitter(color = 'darkblue' , alpha = 0.3) + labs(x='Actual Values', y='Fitted Values') +
        geom_smooth(method = 'lm' , se = FALSE , color = 'black' , size = 0.5) + 
        annotate('text' , x = 90 , y = 25 , label = 'Correlation: 0.71') + 
        geom_abline(slope = 1 , linetype = 'dotted' , color = 'darkblue' , size = 0.5)
```

8) Residuals and Actual values
```{r}
ggplot(data.frame(actual = final_set$death_rate , res = model$residuals) , aes(x = actual , y = res)) + 
        geom_jitter(color = 'darkred' , alpha = 0.3) + labs(x='Actual Values', y='Residuals') + 
        geom_hline(yintercept = 0 , linetype = 'dotted' , size = 1)
```

E. Improvements
Improvements to the model should focus on accounting for outliers and adding additional variables to the model.